Naive Bayes Classifier

### 2019-02-25 hiding/NB_2r2c.Rmd

---
title: "Naive Bayes Analysis Ver 2r.2c"
author: "Dan Johnson"
date: "2/21/2019"
output: html_document
---

@@

##Summary
This it the current Naive Bayes workflow used to classify TA comments based on Subject and Structure. Additions and revisions made to **Version 2r.2c** are:

*  Random selection of data rows used for training and test datasets. 
*  Final output matrix tables are more informative. 
*  Addition of a back-analysis step that reviews entire dataset, not just testing sub-set.
*  Outputs a tidy data table in CSV format that contains the original comments (pre-clean-up) plus their hand-scored and predicted subject and structure.

@@

##Initial Setup
Calling the required libraries.

```{r libraries}
library(tidyverse)
library(tm)
library(tidytext)
library(e1071)
library(caret)
```

##Version 2.0.r of Naive Bayes (NB) Protocol
###Initial Data Input
```{r input data}
#Read in TA comments from CSV file.
base_dataVERB <- read_csv(file='data/coded_full_comments_dataset_Spring18anonVERB.csv')
```

###Extract, Organize Data Subset
Extract comments and classification data as subsets of CSV.
```{r subset data}
#Select rows representing the sub-groups to compare. 
comments_subsetVERB <- filter(base_dataVERB,code.subject=="1_basic"|code.subject=="2_writing"|code.subject=="3_technical"|code.subject=="4_logic")

@@

```{r randomizer}
#Generates a reproducible vector of ~10,000 randomly assorted numbers using set.seed 
#Adjust the number so it matches the number of rows in the original data frame
#Next, set the seed for random numbers, and randomize the order of the vector
vector <- 1:9340 #second value must match number of rows in working dataset
set.seed(123) #change this seed value, and the data are randomized in a new sequence.
vector <- sample(vector)
glimpse(vector)

@@

```{r corpusmaker}
#Convert data set to a volative corpus using "tm" library.
#Create the volatile corpus that contains the "text" vector from data frame.
comments_corpusVERB_R <- VCorpus(VectorSource(comments_rawVERB_R$text))
print(comments_corpusVERB_R)

@@

###Text Data Transformations
The text data transforms are separated out as individual "tm_map"" and "content_transformer" functions to simplify the process of testing various combinations. Conversion to lower case and removing punctuation are the baseline set. Switch between "{r}" and "{}" to turn the other transformation blocks off or on. Removing extra white spaces must be the last transformation, and is required.

@@

#The option to apply stemming has been removed from this version.

@@

#Look at an example of cleaned text comment to see if it fits what is expected.
as.character((comments_corpus_cleanVERB_R[[5]]))
```

@@

##Basic Naive Bayes (NB) Method Using N-grams
This code block uses tm's NLP commands. The same block can be used to assess 1-, 2-, or 3-grams by modifying the "#" variable in "(words(x), #)".

@@

###Data Splitting
Split data into 75% training and 25% testing sets, so that after Naive Bayes spam filter algorithm is built it can be applied to unseen data.

@@

Saves 2 vectors that label the rows in the training and testing vectors
```{r labeler}
comments_all_labels_subject <- comments_rawVERB_R[1:9340,]$subject
comments_train_labels_subject <- comments_rawVERB_R[1:7005, ]$subject
comments_test_labels_subject <- comments_rawVERB_R[7006:9340,]$subject

@@

The proportion of each sub-category should be similar (<1% difference) between the full dataset, and training and testing data subsets.
```{r proportion_check}
prop.table(table(comments_all_labels_subject))
prop.table(table(comments_train_labels_subject))
prop.table(table(comments_test_labels_subject))

@@

###Preparation for Naive Bayes
Remove words from the matrix that appear less than 5 times.
```{r common_words}
comments_freq_words <- findFreqTerms(comments_dtm_train, 5)
str(comments_freq_words)
```

Limit Document Term Matrix to only include words in the comments_freq_words vector. Using all the rows, but we want to limit the columns to these words in the frequency vector.
```{r limit_dtm}
comments_dtm_freq_all <- comments_dtm_all[ , comments_freq_words]
comments_dtm_freq_train <- comments_dtm_train[ , comments_freq_words]
comments_dtm_freq_test <- comments_dtm_test[ , comments_freq_words]
```

The basic e1071 naive bayes classifier works with categorical features, so the matrix must be converted "yes" and "no" categorical variables. This is done using a **convert_counts function** and applying it to the data. This replaces values greater than 0 with yes, and values not greater than 0 with no. 
```{r convert_function}
convert_counts2 <- function(x) {
  x <- ifelse(x > 0, "Yes", "No")
}
```

@@

```{}
# Alternate version of convert_function code block. This version KEEPs word frequencies while still using the same variable names and code structure. The resulting matrices have cells with word frequencies instead. 
#Only works when the NB classifier is looking at ingroup/outgroup. Breaks on multiple categories.
#
convert_counts2 <- function(x) {
  y <- x
}
```

```{r apply_conversion}
comments_all <- apply(comments_dtm_freq_all, MARGIN = 2, convert_counts2)
comments_train <- apply(comments_dtm_freq_train, MARGIN = 2, convert_counts2)
comments_test <- apply(comments_dtm_freq_test, MARGIN = 2, convert_counts2)
```

@@

###Train Model, Predict, Evaluate for Subject
This block uses the e1071 package to implement Naive Bayes algorithm on the data, and predict whether a message is likely to be in group 1_basic, 2_writing, 3_technical, or 4_logic. It evaluates the prediction with the actual data using a crosstable from the gmodels package.
```{r bayes_training}
# Train the classifier on training data, then test on test subset and full dataset.
comments_classifier_subject <- naiveBayes(comments_train, comments_train_labels_subject, laplace=1)
comments_test_pred_subject <- predict(comments_classifier_subject, comments_test)
comments_all_pred_subject <- predict(comments_classifier_subject, comments_all)

comments_classifier_structure <- naiveBayes(comments_train, comments_train_labels_structure, laplace=1)
comments_test_pred_structure <- predict(comments_classifier_structure, comments_test)
comments_all_pred_structure <- predict(comments_classifier_structure, comments_all)

# Create a truth table by tabulating the predicted class labels with the actual class labels 
table("Predictions"= comments_test_pred_subject,  "Actual Subject Test" = comments_test_labels_subject )
table("Predictions"= comments_all_pred_subject,  "Actual Subject All" = comments_all_labels_subject )
table("Predictions"= comments_test_pred_structure,  "Actual Structure Test" = comments_test_labels_structure )
table("Predictions"= comments_all_pred_structure,  "Actual Structure All" = comments_all_labels_structure )

#The truth table lists how many of the predicted categories are in their correct categories. Adding diagonally and dividing by total count gives the percent accuracy at predicting subject. this is what the confusion matrix calculates.

#The faster way is to calculate a confusion matrix, which includes BOTH the truth table, and the final overall accuracy.

#Alternatively, the Cross Table command below (from "gmodels" package) generates a more detailed output table with frequency values for each individual cell.
#CrossTable(comments_test_pred_subject, comments_test_labels_subject, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
```

```{r reportOut}
#"line" is just a spacer text for readability.
line<-("**************************************************************")

@@

# This print-out has both the overall and by-group accuracy values.
conf.mat1
print (line)
print ("Test, Subject")
# This print-out has by-group accuracy values in a single column. 
conf.mat1$byClass
print (line)
print ("Test, Subject")

# This print-out is a single line table with overall accuracy, kappa, lower and higher limits of accuracy, accuracy null value, P value for accuracy, and Mcnemar P value.
conf.mat1$overall
print (line)
print ("Test, Subject")

# This outputs JUST the overall accuracy value. 
conf.mat1$overall['Accuracy']
print ("Test, Subject")

# This print-out has both the overall and by-group accuracy values.
conf.mat2
print (line)
print ("All, Subject")

# This print-out has by-group accuracy values in a single column. 
conf.mat2$byClass
print (line)
print ("All, Subject")

# This print-out is a single line table with overall accuracy, kappa, lower and higher limits of accuracy, accuracy null value, P value for accuracy, and Mcnemar P value.
conf.mat2$overall
print (line)
print ("All, Subject")

# This outputs JUST the overall accuracy value. 
conf.mat2$overall['Accuracy']
print ("All, Subject")

# This print-out has both the overall and by-group accuracy values.
conf.mat3
print (line)
print ("Test, Structure")

# This print-out has by-group accuracy values in a single column. 
conf.mat3$byClass
print (line)
print ("Test, Structure")

# This print-out is a single line table with overall accuracy, kappa, lower and higher limits of accuracy, accuracy null value, P value for accuracy, and Mcnemar P value.
conf.mat3$overall
print (line)
print ("Test, Structure")

# This outputs JUST the overall accuracy value. 
conf.mat3$overall['Accuracy']
print ("Test, Structure")

# This print-out has both the overall and by-group accuracy values.
conf.mat4
print (line)
print ("All, Structure")

# This print-out has by-group accuracy values in a single column. 
conf.mat4$byClass
print (line)
print ("All, Structure")

# This print-out is a single line table with overall accuracy, kappa, lower and higher limits of accuracy, accuracy null value, P value for accuracy, and Mcnemar P value.
conf.mat4$overall
print (line)
print ("All, Structure")

# This outputs JUST the overall accuracy value. 
print ("All, Structure")
conf.mat4$overall['Accuracy']

```

@@

###Output for Review, Analysis
The code block below creates a CSV file containing the original TA comments, the subject and structure labels assigned by a human observer, and the corresponding predicted categories for each comment based on the Naive Bayes estimator. Results are exported to a CSV file. By default the CSV file should be saved to the current working directory.

@@

***
***

This is an alternate method to get the same data tables out using ```tidy``` format. It seems much longer to me, so likely is over-written.
```{r tidy_prediction_table}
#Use broom::tidy to reorganize into a data frame.
tidy_tally1 <-tidy(final.tally)
tidy_tally2 <-tidy(final.tally2)
tidy_tally3 <-tidy(final.tally)
tidy_tally4 <-tidy(final.tally2)

@@

#Spread tidy data into a readable table
spread(prop_tally1,key=Hand.Coded.Structure, value=prop)
spread(prop_tally2,key=Bayes.Predicted.Structure, value=prop)
spread(prop_tally3,key=Hand.Coded.Subject, value=prop)
spread(prop_tally4,key=Bayes.Predicted.Subject, value=prop)
```

```{r}
#Calculate the sum of values for each group.
proptallysum1<-prop_tally1 %>% group_by(Hand.Coded.Subject) %>% summarize(total=sum(prop))
proptallysum1
proptallysum2<-prop_tally2 %>% group_by(Bayes.Predicted.Subject) %>% summarize(total=sum(prop))
proptallysum2
proptallysum3<-prop_tally3 %>% group_by(Hand.Coded.Structure) %>% summarize(total=sum(prop))
proptallysum3
proptallysum4<-prop_tally4 %>% group_by(Bayes.Predicted.Structure) %>% summarize(total=sum(prop))
proptallysum4

```

@@

Need to combine two datasets into one.

>>>>

---
title: "Naive Bayes Classifier"
author: "Dan Johnson"
date: "2/21/2019"
output: html_document
---

@@

#Overview
This is the current version of the Naive Bayes workflow used to classify TA comments based on Subject and Structure. Features of **Version 2r.2c**:

*  Data rows used for training and test datasets are assigned randomly but reproducibly using 'set.seed'. 
*  Detailed final output matrix tables. 
*  Back-analysis evaluates both a testing sub-set and the entire input dataset.
*  Outputs a tidy data table in CSV format with unmodified original comments, and hand-scored and NB-predicted Subject and Structure assignments.

This page outlines Ver 2r.2c of NB Classifier and the process for building various components used in the project. Details of each  component are provided in these sub-pages.

For structure of the initial dataset see: https://adanieljohnson.github.io/default_website/datastructure.html

For import and post-import data validation see: https://adanieljohnson.github.io/default_website/dataimport.html

@@

#Initial Setup
Calling required libraries.
```{r libraries}
library(tidyverse)
library(tm)
library(tidytext)
library(e1071)
library(caret)
```

Import the initial dataset. Then extract comments and classification data as subsets of CSV.
```{r input data}
#Read in TA comments from CSV file.
base_dataVERB <- read_csv(file='data/coded_full_comments_dataset_Spring18anonVERB.csv')

#Select rows representing the sub-groups to compare. 
comments_subsetVERB <- filter(base_dataVERB,code.subject=="1_basic"|code.subject=="2_writing"|code.subject=="3_technical"|code.subject=="4_logic")

@@

Randomize data used for training and testing sets. This block of code generates a reproducible vector of ~10,000 randomly assorted numbers using set.seed, then writes the vector back to data table.
```{r randomizer}
#Adjust the number so it matches the number of rows in the original data frame
#Next, set the seed for random numbers, and randomize the order of the vector
vector <- 1:9340 #second value must match number of rows in working dataset
set.seed(123) #change this seed value, and the data are randomized in a new sequence.
vector <- sample(vector)
glimpse(vector)

@@

Convert data set to a volative corpus using "tm" library.
```{r corpusmaker}
#Create the volatile corpus that contains the "text" vector from data frame.
comments_corpusVERB_R <- VCorpus(VectorSource(comments_rawVERB_R$text))
print(comments_corpusVERB_R)

@@

#Text Data Transformations
The text data transforms in this code block are intentially separate "tm_map"" and "content_transformer" functions. This simplifies the process of testing how various combinations affect classification. Converting the text to lower case and removing punctuation are the baseline transformations. Switch between "{r}" and "{}" to turn other transformation blocks off or on as desired. **Removing extra white spaces must be the last transformation, and is required.**

@@

#The option to apply stemming has been removed from this version. Stemming reduced accuracy in all trials. Future iterations should examine lemmatization instead.

@@

#Look at an example of cleaned text comment to see if cleaned comment matches what is expected. Change the value inside double brackets to look at a different comment.
as.character((comments_corpus_cleanVERB_R[[5]]))
```

@@

#Tokenize the N-Grams
This code block uses tm's NLP commands to assess different n-grams. For single words, change the "#" variable in "(words(x), #)" to "1". Use other values to change the n-gram size. Data are stored in document-term matrix (dtm) format.

@@

#Split Data to Training, Testing Sets
This code splits data into 75% training and 25% testing sets, so that after Naive Bayes algorithm is trained it can be tested on unseen data.

@@

This saves 3 vectors containing labels for the rows in the training and testing vectors
```{r labeler}
comments_all_labels_subject <- comments_rawVERB_R[1:9340,]$subject
comments_train_labels_subject <- comments_rawVERB_R[1:7005, ]$subject
comments_test_labels_subject <- comments_rawVERB_R[7006:9340,]$subject

@@

This checks the proportions of sub-categories in training and testing groups. The proportion of each sub-category should be similar (<1% difference) between the full dataset, and training and testing data subsets.
```{r proportion_check}
prop.table(table(comments_all_labels_subject))
prop.table(table(comments_train_labels_subject))
prop.table(table(comments_test_labels_subject))

@@

#Preparing Data for Naive Bayes Training
Remove words from the document-term matrix that appear less than 5 times.
```{r common_words}
comments_freq_words <- findFreqTerms(comments_dtm_train, 5)
str(comments_freq_words)
```

Limit document-term matrix to words in the comments_freq_words vector. We are using all of the rows, but we want to limit the columns to these words in the frequency vector.
```{r limit_dtm}
comments_dtm_freq_all <- comments_dtm_all[ , comments_freq_words]
comments_dtm_freq_train <- comments_dtm_train[ , comments_freq_words]
comments_dtm_freq_test <- comments_dtm_test[ , comments_freq_words]
```

The e1071 Naive Bayes classifier works with categorical features, so the matrix must be converted "yes" and "no" categorical variables. This is done using a **convert_counts function** and applying it to the data. This replaces values greater than 0 with yes, and values not greater than 0 with no. 
```{r convert_function}
convert_counts2 <- function(x) {
  x <- ifelse(x > 0, "Yes", "No")
}
```

@@

Below is an **alternate version** of convert_function code block that KEEPs word frequencies while still using the same variable names and code structure. The processed matrices have cells with word frequencies instead of categorical yes/no values. This block only works for binary classification (ingroup/outgroup). It breaks when applied to data with multiple categories.
```{}
convert_counts2 <- function(x) {
  y <- x
}
```

Applies whichever "convert_counts2" function is currently active.
```{r apply_conversion}
comments_all <- apply(comments_dtm_freq_all, MARGIN = 2, convert_counts2)
comments_train <- apply(comments_dtm_freq_train, MARGIN = 2, convert_counts2)
comments_test <- apply(comments_dtm_freq_test, MARGIN = 2, convert_counts2)
```

@@

#Train Model, Predict, Evaluate for Subject
This block predicts whether a message is likely to be in group 1_basic, 2_writing, 3_technical, or 4_logic. 
```{r bayes_training}
# Train the classifier on training data, then test on test dataset, and full dataset.
comments_classifier_subject <- naiveBayes(comments_train, comments_train_labels_subject, laplace=1)
comments_test_pred_subject <- predict(comments_classifier_subject, comments_test)
comments_all_pred_subject <- predict(comments_classifier_subject, comments_all)

comments_classifier_structure <- naiveBayes(comments_train, comments_train_labels_structure, laplace=1)
comments_test_pred_structure <- predict(comments_classifier_structure, comments_test)
comments_all_pred_structure <- predict(comments_classifier_structure, comments_all)
```

A truth table lists how many of the predicted categories are in their correct categories. Adding diagonally and dividing by total count gives the percent accuracy at predicting Subject. 
```{r}
# Create a truth table by tabulating the predicted class labels with the actual class labels 
table("Predictions"= comments_test_pred_subject,  "Actual Subject Test" = comments_test_labels_subject )
table("Predictions"= comments_all_pred_subject,  "Actual Subject All" = comments_all_labels_subject )
table("Predictions"= comments_test_pred_structure,  "Actual Structure Test" = comments_test_labels_structure )
table("Predictions"= comments_all_pred_structure,  "Actual Structure All" = comments_all_labels_structure )

#An alternative is to use the Cross Table command (from "gmodels" package) to generate a more detailed output table with frequency values for each individual cell.
#CrossTable(comments_test_pred_subject, comments_test_labels_subject, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
```

The faster method is to calculate a confusion matrix, which displays BOTH the truth table, and the final overall accuracy. A confusion matrix can be queried for specific values.
```{r reportOut}
#"line" is just spacer text for improving readability.
line<-("**************************************************************")

@@

#These commands show how to print out individual pieces of the confusion matrix. They show both the overall and by-group accuracy values for Subject x Test Subset.
conf.mat1
print (line)
print ("Test, Subject")

# This block outputs the by-group accuracy values in a single column for Subject x Test Subset. 
conf.mat1$byClass
print (line)
print ("Test, Subject")

# This block outputs a single line table with overall accuracy, kappa, lower and higher limits of accuracy, accuracy null value, P value for accuracy, and Mcnemar P value for Subject x Test Subset.
conf.mat1$overall
print (line)
print ("Test, Subject")

# This block outputs JUST the overall accuracy value for Subject x Test Subset. 
conf.mat1$overall['Accuracy']
print ("Test, Subject")

# This block outputs both the overall and by-group accuracy values for Subject x All Data.
conf.mat2
print (line)
print ("All, Subject")

# This block outputs by-group accuracy values in a single column for Subject x All Data. 
conf.mat2$byClass
print (line)
print ("All, Subject")

# This block outputs a single line table with overall accuracy, kappa, lower and higher limits of accuracy, accuracy null value, P value for accuracy, and Mcnemar P value for Subject x All Data.
conf.mat2$overall
print (line)
print ("All, Subject")

# This block outputs JUST the overall accuracy value for Subject x All Data. 
conf.mat2$overall['Accuracy']
print ("All, Subject")

# This block outputs both the overall and by-group accuracy values for Structure x Test Subset.
conf.mat3
print (line)
print ("Test, Structure")

# This block outputs by-group accuracy values in a single column for Structure x Test Subset. 
conf.mat3$byClass
print (line)
print ("Test, Structure")

# This block outputs a single line table with overall accuracy, kappa, lower and higher limits of accuracy, accuracy null value, P value for accuracy, and Mcnemar P value for Structure x Test Subset.
conf.mat3$overall
print (line)
print ("Test, Structure")

# This block outputs JUST the overall accuracy value for Structure x Test Subset. 
conf.mat3$overall['Accuracy']
print ("Test, Structure")

# This block outputs both the overall and by-group accuracy values for Structure x All Data.
conf.mat4
print (line)
print ("All, Structure")

# This block outputs by-group accuracy values in a single column for Structure x All Data. 
conf.mat4$byClass
print (line)
print ("All, Structure")

# This block outputs a single line table with overall accuracy, kappa, lower and higher limits of accuracy, accuracy null value, P value for accuracy, and Mcnemar P value for Structure x All Data.
conf.mat4$overall
print (line)
print ("All, Structure")

# This block outputs JUST the overall accuracy value for Structure x All Data. 
print ("All, Structure")
conf.mat4$overall['Accuracy']
```

@@

#Output for Review, Analysis
The code block below creates a CSV file containing the original TA comments, the subject and structure labels assigned by a human observer, and the corresponding predicted categories for each comment based on the Naive Bayes estimator. Results are exported to a CSV file. By default the CSV file should be saved to the current working directory.

@@

This is an alternate method to get the same data tables out using ```tidy``` format. It seems much longer to me, so likely is over-written.
```{}
tidy_prediction_table

#Use broom::tidy to reorganize into a data frame.
tidy_tally1 <-tidy(final.tally)
tidy_tally2 <-tidy(final.tally2)
tidy_tally3 <-tidy(final.tally)
tidy_tally4 <-tidy(final.tally2)

@@

#Spread tidy data into a readable table
spread(prop_tally1,key=Hand.Coded.Structure, value=prop)
spread(prop_tally2,key=Bayes.Predicted.Structure, value=prop)
spread(prop_tally3,key=Hand.Coded.Subject, value=prop)
spread(prop_tally4,key=Bayes.Predicted.Subject, value=prop)

#Calculate the sum of values for each group.
proptallysum1<-prop_tally1 %>% group_by(Hand.Coded.Subject) %>% summarize(total=sum(prop))
proptallysum1
proptallysum2<-prop_tally2 %>% group_by(Bayes.Predicted.Subject) %>% summarize(total=sum(prop))
proptallysum2
proptallysum3<-prop_tally3 %>% group_by(Hand.Coded.Structure) %>% summarize(total=sum(prop))
proptallysum3
proptallysum4<-prop_tally4 %>% group_by(Bayes.Predicted.Structure) %>% summarize(total=sum(prop))
proptallysum4
```

@@

***
Need to combine two datasets into one. 
***
